Clustering time series based on probability distributions across temporal granularities

### 2021-11-12 main.Rmd

**JS-based distances**

@@

**wpd-based distances**

@@

_Number of clusters_: Different approaches to determining cluster size are required for varied applications and research objectives. Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired primarily by defining and balancing "within-cluster homogeneity" and "between-cluster separation." Cluster separation index ($sindex$) proposed by @Hennig2014-ah is employed in this paper to have less dependency on a single or few ambiguous points. $sindex$ [@R-fpc] is determined by calculating the distance between each observation and the cluster to which it does not belong. Then, the least 10% of these distances are averaged to only evaluate places near the cluster edges. $sindex$, like other statistics in the literature, cannot be optimized over the number of clusters, since increasing the number of clusters reduces it. The number of clusters required is determined by the slope going from steep to shallow (an elbow).

_ALgorithm_: With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input. A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy, or more recent approaches like distribution, density, and others. We employ agglomerative hierarchical clustering in conjunction with Ward's linkage. Hierarchical cluster techniques fuse neighboring points sequentially to form bigger clusters, beginning with a full pairwise distance matrix. The distance between clusters is described using a "linkage technique". This agglomerative approach successively merges the pair of clusters with the shortest between-cluster distance using Ward's linkage method.

@@

## Characterization of clusters

@@

## Data generating processes

@@

_Scenario (S1): All significant granularities_

@@

_Scenario (S3): One significant granularity_

@@

## Visual exploration of findings

@@

_Data filtering and variable selection_

@@

_Prototype selection_

>>>>

@@

@@

### Number of clusters

Different approaches to determining cluster size are required for varied applications and research objectives. Some typical ways of determining the number of clusters in the literature (e.g., the gap statistic [@tibshirani2001estimating], average silhouette width [@rousseeuw1987silhouettes], Dunn index [@dunn1973fuzzy]) are inspired primarily by defining and balancing "within-cluster homogeneity" and "between-cluster separation." Cluster separation index ($sindex$) proposed by @Hennig2014-ah is employed in this paper to have less dependency on a single or few ambiguous points. $sindex$ [@R-fpc] is determined by calculating the distance between each observation and the cluster to which it does not belong. Then, the least 10% of these distances are averaged to only evaluate places near the cluster edges. $sindex$, like other statistics in the literature, cannot be optimized over the number of clusters, since increasing the number of clusters reduces it. The number of clusters required is determined by the slope going from steep to shallow (an elbow).

### Algorithm

With a way to obtain pairwise distances, any clustering algorithm can be employed that supports the given distance metric as input. A good comprehensive list of algorithms can be found in @Xu2015-ja based on traditional ways like partition, hierarchy, or more recent approaches like distribution, density, and others. We employ agglomerative hierarchical clustering in conjunction with Ward's linkage. Hierarchical cluster techniques fuse neighboring points sequentially to form bigger clusters, beginning with a full pairwise distance matrix. The distance between clusters is described using a "linkage technique". This agglomerative approach successively merges the pair of clusters with the shortest between-cluster distance using Ward's linkage method.

@@

### Characterization of clusters

@@

## Data generation

@@

_Scenario (S1): All granularities significant_

@@

_Scenario (S3): Only one significant granularity_

@@

## Visual exploration of results

@@

## Data filtering and variable selection

@@

## Prototype selection
